State of the Practice for Lattice Boltzmann Method Software

### 2022-01-27 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Removed extra space in header for Table 5
Finished revision of Section 3 (Ranking Projects)

Figure~\ref{Fig_Installability} shows the installability ranking of the software
packages using AHP. Software packages with a higher score (ESPResSo, MechSys,
OpenLB, DL\_MESO, Palabos) tend to have one set of linear installation
instructions written as if the person doing the installation has none of the
dependencies installed. The instructions often list compatible operating system
versions and include instructions for the installation of dependencies. The top
ranked packages often incorporate some sort of automation of the installation
process and have fewer manual installation steps. The number of dependencies a
package has does not correlate with a higher score. The ability to validate the
installation process, often through tutorials or test examples that include
expected output, is correlated with a higher score. Active work on a project
seems to improve installability, since the top seven ranked packages are all
classified as alive.

@@

Figure~\ref{Fig_Reliability} shows the surface reliability ranking of the
software packages using AHP. Software packages with a high score either did not
break during installation, or the broken installation was recoverable. All of
the top five ranked packages have tutorials. The package pyLBM broke
during tutorial testing, but a descriptive error message helped in recovery.
Nine of the top 10 ranked packages are noted as being alive.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages using AHP. Software packages with a high score behaved reasonably in
response to unexpected input as described above. All of the software packages
that installed correctly passed this test. They output descriptive error
messages or at least do not crash. Software packages with a lower surface
robustness score did not install correctly, so their robustness score may not be
a true reflection of runtime robustness. All successfully installed software
packages that require a plain text input file correctly handled an unexpected
change to these input files, including a replacement of new lines with carriage
returns. Furthermore, nine of the top 10 ranked packages are noted as being
alive.

@@

Figure~\ref{Fig_Usability} shows the surface usability ranking of the software
packages using AHP. Software packages with a high score have a tutorial and user
manual, sometimes have documented user characteristics, and have at least one
user support model. Many packages have several user support models. Furthermore,
four of the top five packages are noted as being alive.

@@

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{2cm}p{2.5cm}p{2cm}p{2.5cm} }
			\toprule
			Name & Text Files & Binary Files & LOC & Avg. LOC / Text File\\
			\midrule
			DL\_MESO (LBE) & 310 & 51 & 170223 & 549\\
			ESPResSo & 1390& 83 & 195083 & 140\\
			ESPResSo++ & 1406& 30 & 165194 & 118\\
			HemeLB & 1102& 44 & 123806 & 112\\
			laboetie & 133& 1 & 48403 & 364\\		
			LatBo.jl & 41& 0 & 42172 & 1029\\
			LB2D-Prime & 82& 19 & 54755 & 668\\
			LB3D & 99 & 76 & 39766 & 402\\
			LB3D-Prime & 23& 6 & 12944 & 563\\
			lbmpy& 250 & 29 & 61632 & 247\\
			lettuce & 73 & 1 & 7660 & 105\\
			LIMBES & 26 & 1 & 4872 & 187\\
			Ludwig & 954 & 32 & 162518 & 170\\
			LUMA & 314 & 19 & 4399723 & 14011\\
			MechSys & 333 & 3 & 95707 & 287\\
			MP-LABS & 307 & 3 & 43124 & 140\\
			Musubi & 1347 & 1839 & 281879 & 209\\
			OpenLB & 1438 & 7 & 218406 & 152\\
			Palabos & 1974 & 71 & 563841 & 286\\
			pyLBM & 272 & 108 & 37234 & 137\\
			Sailfish & 632 & 11 & 69398 & 110\\
			SunlightLB & 36 & 1 & 7646 & 212\\
			TCLB & 594 & 7 & 49156 & 83\\
			waLBerla & 2643 & 69 & 873988 & 331\\
			\bottomrule
		\end{tabular}
		\caption{Module Data} \label{moduledata}
	\end{center}
\end{table}

@@

\subsection{Surface Understandability}

@@

Figure~\ref{Fig_Understandability} shows the surface understandability ranking
of the software packages using AHP. Software packages with a high score have a
consistent indentation and formatting style, and consistent, distinctive, and
meaningful code identifiers. They also have symbolic constants, and explicitly
identify mathematical and LBM algorithms. Their comments are clear and indicate
what is being done in the source code. The source code is well modularized and
structured. Furthermore, all of the top five ranked packages are noted as being
alive.

@@

Software package artifacts were reviewed for the identification of a specific
development model, like a waterfall of agile development model, and the presence
of documentation recording the development process and standard. They were also
reviewed for the identification of the development environment, and the presence
of release notes. The packages tended to not explicitly use well-known
development models. This was also noted in the interviews with developers, as
detailed below. The development teams of these packages are fairly small and
easily organized without the need for such processes. Seven of the software
packages did have some artifacts outlining the general development process, how
to contribute, and the status of the package or its components. Eight of the
packages have artifacts that note the development environment. While this
information could help developers, and would improve transparency, the small
close-knit nature of the development teams make explicitly publicly specifying
this information practically unnecessary. Version release notes were found in
nine of the software packages.

@@

Figure~\ref{Fig_VisibilityTransparency} shows the visibility and transparency
ranking of the software packages using AHP. Software packages with a high score
have an explicit development model and defined development process. They also
had detailed and easy to access notes accompanying software releases.
Furthermore, four of the top five ranked packages are noted as being alive.
MP-LABS is noted as being dead.

\subsection{Overall Quality}

@@

Figure~\ref{Fig_OverallScore} shows the overall ranking of the software packages
using AHP. In the absence of other information on priorities, the overall
ranking was calculated by assuming an equal weighting between all qualities. If the
qualities were weighed differently, the overall software package ranking would
change. Software packages with an overall high score ranked high in at least
several of the individual qualities that were quantitatively measured.

Looking at the top three ranked packages: ESPResSo had achieved a relative high
score in installability, surface usability, maintainability, reusability, and
visibility and transparency. Ludwig scored high in surface correctness and
verifiability, surface robustness, surface usability, maintainability, and
visibility and transparency. Palabos scored high in installability, surface
reliability, surface robustness, maintainability, understandability, and
visibility and transparency.

@@

The assumption that more code files is an indicator of reusability
(Section~\ref{reusabilityresults}) may not be a valid assumption.

>>>>

Figure~\ref{Fig_Installability} shows the installability ranking of the software
packages. Software packages with a higher score (ESPResSo, MechSys, OpenLB,
DL\_MESO, Palabos) tend to have one set of linear installation instructions
written as if the person doing the installation has none of the dependencies
installed. The instructions often list compatible operating system versions and
include instructions for the installation of dependencies. The top ranked
packages often incorporate some sort of automation of the installation process
and have fewer manual installation steps. The number of dependencies a package
has does not correlate with a higher score. The ability to validate the
installation process, often through tutorials or test examples that include
expected output, is correlated with a higher score. Active work on a project
seems to improve installability, since the top seven ranked packages are all
classified as alive.

@@

Figure~\ref{Fig_Reliability} shows the surface reliability ranking of the
software packages. Software packages with a high score either did not break
during installation, or the broken installation was recoverable. All of the top
five ranked packages have tutorials. The package pyLBM broke during tutorial
testing, but a descriptive error message helped in recovery. Nine of the top 10
ranked packages are noted as being alive.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages. Software packages with a high score behaved reasonably in response to
unexpected input as described above. All of the software packages that installed
correctly passed this test. They output descriptive error messages or at least
do not crash. Software packages with a lower surface robustness score did not
install correctly, so their robustness score may not be a true reflection of
runtime robustness. All successfully installed software packages that require a
plain text input file correctly handled an unexpected change to these input
files, including a replacement of new lines with carriage returns. Furthermore,
nine of the top 10 ranked packages are noted as being alive.

@@

Figure~\ref{Fig_Usability} shows the surface usability ranking of the software
packages. Software packages with a high score have a tutorial and user manual,
sometimes have documented user characteristics, and have at least one user
support model. Many packages have several user support models. Furthermore, four
of the top five packages are noted as being alive.

@@

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{2cm}p{2.5cm}p{2cm}p{2.5cm} }
			\toprule
			Name & Text Files & Binary Files & LOC & Avg.\ LOC / Text File\\
			\midrule
			DL\_MESO (LBE) & 310 & 51 & 170223 & 549\\
			ESPResSo & 1390& 83 & 195083 & 140\\
			ESPResSo++ & 1406& 30 & 165194 & 118\\
			HemeLB & 1102& 44 & 123806 & 112\\
			laboetie & 133& 1 & 48403 & 364\\		
			LatBo.jl & 41& 0 & 42172 & 1029\\
			LB2D-Prime & 82& 19 & 54755 & 668\\
			LB3D & 99 & 76 & 39766 & 402\\
			LB3D-Prime & 23& 6 & 12944 & 563\\
			lbmpy& 250 & 29 & 61632 & 247\\
			lettuce & 73 & 1 & 7660 & 105\\
			LIMBES & 26 & 1 & 4872 & 187\\
			Ludwig & 954 & 32 & 162518 & 170\\
			LUMA & 314 & 19 & 4399723 & 14011\\
			MechSys & 333 & 3 & 95707 & 287\\
			MP-LABS & 307 & 3 & 43124 & 140\\
			Musubi & 1347 & 1839 & 281879 & 209\\
			OpenLB & 1438 & 7 & 218406 & 152\\
			Palabos & 1974 & 71 & 563841 & 286\\
			pyLBM & 272 & 108 & 37234 & 137\\
			Sailfish & 632 & 11 & 69398 & 110\\
			SunlightLB & 36 & 1 & 7646 & 212\\
			TCLB & 594 & 7 & 49156 & 83\\
			waLBerla & 2643 & 69 & 873988 & 331\\
			\bottomrule
		\end{tabular}
		\caption{Module Data} \label{moduledata}
	\end{center}
\end{table}

@@

\subsection{Surface Understandability} \label{Sec_SurfUnderstandability}

@@

Figure~\ref{Fig_Understandability} shows the surface understandability ranking
of the software packages. Software packages with a high score have a consistent
indentation and formatting style, and consistent, distinctive, and meaningful
code identifiers. They also have symbolic constants, and explicitly identify
mathematical and LBM algorithms. Their comments are clear and indicate what is
being done in the source code. The source code is well modularized and
structured. All of the top five ranked packages are noted as being alive.

@@

Software package artifacts were reviewed for the identification of a specific
development model, like a waterfall of agile development model, and the presence
of documentation recording the development process and standard. They were also
reviewed for the identification of the development environment, and the presence
of release notes. The packages tended to not explicitly use well-known
development models. This was also noted in the interviews with developers
(Section~\ref{Sec_CompareMethodologies}). The development teams of these
packages are fairly small and easily organized without the need for such
processes. Seven of the software packages did have some artifacts outlining the
general development process, how to contribute, and the status of the package or
its components. Eight of the packages have artifacts that note the development
environment. While this information could help developers, and would improve
transparency, the small close-knit nature of the development teams make
explicitly publicly specifying this information practically unnecessary. Version
release notes were found in nine of the software packages.

@@

Figure~\ref{Fig_VisibilityTransparency} shows the visibility and transparency
ranking of the software packages. Software packages with a high score have an
explicit development model and defined development process. They also had
detailed and easy to access notes accompanying software releases. Four of the
top five ranked packages are noted as being alive.

\subsection{Overall Quality} \label{Sec_OverallQuality}

@@

Figure~\ref{Fig_OverallScore} shows the overall ranking of the software
packages. In the absence of other information on priorities, the overall ranking
was calculated by assuming an equal weighting between all qualities. If the
qualities were weighed differently, the overall software package ranking would
change. Software packages with an overall high score ranked high in at least
several of the individual qualities that were quantitatively measured.

Looking at the top three ranked packages: ESPResSo achieved a relatively high
score in installability, surface usability, maintainability, reusability, and
visibility and transparency. Ludwig scored high in surface correctness and
verifiability, surface robustness, surface usability, maintainability, and
visibility and transparency. Palabos scored high in installability, surface
reliability, surface robustness, maintainability, understandability, and
visibility and transparency.

@@

Problems with the measurement of qualities:

\begin{itemize}
\item The assumption that more code files is an indicator of reusability
(Section~\ref{reusabilityresults}) may not be a valid assumption.
\item Understandability was measured using only 10 random files
(Section~\ref{Sec_SurfUnderstandability}).  This could be improved.
\end{itemize}

Overall ranking with equally weighted qualities isn't realistic (Section~\ref{Sec_OverallQuality}).
